---
title: "Modelling Bernoulli distribution"
author: "Pärt Prommik & Ülo Maiväli"
date: '2022-23-02'
output:
  html_document:
---


# Preparation

- Clear workspace

- Be sure that correct project is opened

- Install necessary packages

```{r}
install.packages("bayestestR")
```

- Define knitr  settings

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 3, fig.height = 2)
```

- make a folder "models" into your project folder

# Load necessary packages

```{r}
library(tidyverse)
library(brms)
library(patchwork)
library(see)
library(bayestestR)
options(mc.cores = parallel::detectCores()) #sets your system to use multiple cores simultaneously while fitting models 
```


# Lets' define a prior in log-odds scale

Let's say we believe that most students are female who like cats. C seems resonable?

```{r, fig.width=6, fig.height=4}
#DEFINE YOUR BELIEF
our_belief = 0.75 #define mean: what proportion like cats (reference level of our Y-variable)

#let's define few standard deviation values for for the mean in log-odds scale (I know, this is not intuitive at first)
sd1 = 1.2 
sd2 = 0.4
sd3 = 0.8
sd4 = 2.5

#PLOT CODE (no need to edit)
prior_data = bind_rows(
  tibble(log_odds = rnorm(1000, logit_scaled(our_belief), sd1), prior = str_c("A: values for log-odds prior, mean = ", round(logit_scaled(our_belief), digits = 2), ", sd = ", sd1) %>% str_wrap(width = 30)),
  tibble(log_odds = rnorm(1000, logit_scaled(our_belief), sd2), prior = str_c("B: values for log-odds prior, mean = ", round(logit_scaled(our_belief), digits = 2), ", sd = ", sd2) %>% str_wrap(width = 30)),
  tibble(log_odds = rnorm(1000, logit_scaled(our_belief), sd3), prior = str_c("C: values for log-odds prior, mean = ", round(logit_scaled(our_belief), digits = 2), ", sd = ", sd3) %>% str_wrap(width = 30)),
  tibble(log_odds = rnorm(1000, logit_scaled(our_belief), sd4), prior = str_c("D: values for log-odds prior, mean = ", round(logit_scaled(our_belief), digits = 2), ", sd = ", sd4) %>% str_wrap(width = 30))) %>% mutate(probability = inv_logit_scaled(log_odds), prior = as_factor(prior))

prior_data %>% ggplot(aes(x = probability))+
  geom_density(fill = "grey80")+
  scale_x_continuous(limits = c(0,1), labels = c(0, "", 0.5, "", 1))+
  theme_classic()+
  facet_wrap(~prior, nrow = 2, ncol = 2, scales = "free")
```

# Let's simulate data

Your answers

```{r}
data = tibble(
  answer = c("Dog-person", "Cat-person", "Cat-person", "Cat-person", "Cat-person", "Cat-person", "Cat-person", "Cat-person", "Cat-person", "Cat-person", "Dog-person", "Dog-person", "Dog-person", "Dog-person", "Dog-person", "Dog-person", "Dog-person", "Dog-person", "Dog-person", "Dog-person", "Dog-person", "Dog-person", "Dog-person", "Dog-person", "Dog-person", "Dog-person"),
  sex = c("Male", "Male", "Male", "Male", "Male", "Female", "Female", "Female", "Female", "Female", "Male", "Male", "Male", "Male", "Female", "Female", "Female", "Female", "Female", "Male", "Male", "Male", "Male", "Male", "Male", "Male")) %>% 
  mutate(answer = as_factor(answer), answer = fct_relevel(answer, "Dog-person", "Cat-person"))
```


View your data

```{r}
data
```

How factor levels are ordered

```{r}
levels(data$answer)
```

Calculate mean and SD

```{r}
summary_table = data %>%
  group_by(answer) %>% 
  summarise(n = n()) %>%
  mutate(proportion = round(n/sum(n), digits = 2))

summary_table
```

Let's make a density plot

It includes only 26 observations, thus it is quite normal, flat (wide) distribution.

```{r}
plot2 = data %>% 
  ggplot(aes(x = 1, fill = answer)) + 
  geom_bar(position = "fill")

plot2
```

# Intercept-only model

## Interceot-only and prior-only model

First, let's see what priors can be defined

```{r}
possible_prior_options = get_prior(
  formula = answer ~ 1,
  data = data,
  family = bernoulli())

possible_prior_options %>% as_tibble()
```

Let's define a prior in log-odds scale

```{r}
our_prior = c(
  prior(normal(1.1,0.8), class = Intercept))

our_prior %>% as_tibble()
```

Let's run a prior only model (the data, likelihood, is not included into calculations)

```{r}
m5.1 = brm(formula = answer ~ 1, 
         data = data, 
         family = bernoulli(), 
         file = "m5.1", 
         seed = 123,
         prior = our_prior,
         sample_prior = "only")
```


- model summary shows that logit link was used for mu

```{r}
m5.1
```


-- to transform intercept to a probability scale, we have to use inv_logit_scaled() function

```{r}
inv_logit_scaled(1.06)
```

- this is what prior-only thinks about the mean proportion of students who are cat persons

```{r}
posterior1 = m5.1 %>% 
  as_draws_df() %>% 
  mutate(b_Intercept_transformed = inv_logit_scaled(b_Intercept))
  
posterior1 %>% summarise(mean_b_Intercept = mean(b_Intercept),
                         mean_b_Intercept_transformed = mean(b_Intercept_transformed))
```


-- let's plot these posterior variables

```{r}
hdi1 = plot(hdi(posterior1$b_Intercept, ci = 0.95)) + ggtitle("Log-odds scale")

hdi2 = plot(hdi(posterior1$b_Intercept_transformed, ci = 0.95)) + ggtitle("Probability scale")

hdi1/hdi2
```

## Intercept-only and default prior model

```{r}
m5.2 = brm(formula = answer ~ 1, 
         data = data, 
         family = bernoulli(), 
         file = "m5.2", 
         seed = 123,
         sample_prior = "yes")
```

- model summary

```{r}
m5.2
```



-- to transform intercept to a probability scale, we have to use inv_logit_scaled() function

```{r}
inv_logit_scaled(-0.48)
```


- this is what our mostly data-based model thinks about the mean proportion of students who are cat persons

```{r}
posterior2 = m5.2 %>% 
  as_draws_df() %>% 
  mutate(b_Intercept_transformed = inv_logit_scaled(b_Intercept),
         prior_Intercept_transformed = inv_logit_scaled(prior_Intercept))
  
hdi3 = plot(hdi(posterior2$b_Intercept_transformed, ci = 0.95)) + ggtitle("Probability scale b_Intercept")

hdi4 = plot(hdi(posterior2$prior_Intercept_transformed, ci = 0.95)) + ggtitle("Used prior in probability scale")

hdi3 / hdi4
```


## Intercept-only model with the informative prior

```{r}
m5.3 = brm(formula = answer ~ 1, 
         data = data, 
         family = bernoulli(), 
         file = "m5.3", 
         seed = 123, 
         prior = our_prior,
         sample_prior = "yes")
```


Model summary
```{r}
m5.3
```





-- to transform intercept to a probability scale, we have to use inv_logit_scaled() function
```{r}
inv_logit_scaled(-0.16)
```


- this is what our data- and prior-based model thinks about the mean proportion of students who are cat persons

```{r}
posterior3 = m5.3 %>% 
  as_draws_df() %>% 
  mutate(b_Intercept_transformed = inv_logit_scaled(b_Intercept),
         prior_Intercept_transformed = inv_logit_scaled(prior_Intercept))
  
hdi5 = plot(hdi(posterior3$b_Intercept_transformed, ci = 0.95)) + ggtitle("Probability scale b_Intercept")

hdi6 = plot(hdi(posterior3$prior_Intercept_transformed, ci = 0.95)) + ggtitle("Used prior in probability scale")

hdi5 / hdi6
```

Let's calculate model summary statistics

```{r}
posterior3 %>% select(b_Intercept, b_Intercept_transformed) %>% 
  posterior_summary() %>% 
  as_tibble(rownames = "Posterior variable") %>% 
  mutate(across(where(is.numeric), round, 2))
```


## Visualize the influence of our prior

```{r, fig.width=7, fig.height=5K}
hdi3 = hdi3 + ggtitle("m4.2: b_Intercept") + scale_x_continuous(limits = c(0,1))
hdi4 = hdi4 + ggtitle("m4.2: default prior") + scale_x_continuous(limits = c(0,1))
hdi5 = hdi5 + ggtitle("m4.3: b_Intercept") + scale_x_continuous(limits = c(0,1))
hdi6 = hdi6 + ggtitle("m4.3: our prior") + scale_x_continuous(limits = c(0,1))

(hdi3 + hdi4)/
(hdi5 + hdi6) + plot_layout(guides = "collect")
```

# Modelling a comparison between 2 groups

Before, students tend to equally like cats and dogs. But are there sex-specific differences?

Let's use the same simulated data

```{r}
data
```

Let's make a summary table

```{r}
summary_table2 = data %>%
  group_by(sex, answer) %>% 
  summarise(n = n()) %>%
  mutate(proportion = round(n/sum(n), digits = 2))

summary_table2
```


Let's plot our data

```{r}
plot2 = data %>% 
  ggplot(aes(x = answer, fill = sex)) + 
  geom_bar(position = "fill")

plot2
```

Chi-square test

```{r}
test = chisq.test(table(data$answer, data$sex))
test
```


Which priors can be set for our new model

```{r}
possible_prior_options2 = get_prior(
  formula = answer ~ sex,
  data = data,
  family = bernoulli())

possible_prior_options2 %>% as_tibble()
```

Let's visualised our beliefs

Let's use weakly-informative priors:
- students may equally like cats or dogs
- female and male students equally like cats or dogs

```{r}
#DEFINE YOUR BELIEF
our_belief2 = 0.5 #define mean: what proportion like cats (reference level of our Y-variable)

#let's define few standard deviation values for for the mean in log-odds scale (I know, this is not intuitive at first)
sd1 = 3
sd2 = 3
sd3 = 1
sd4 = 1

#PLOT CODE (no need to edit)
prior_data = bind_rows(
  tibble(log_odds = rnorm(1000, logit_scaled(our_belief2), sd1), prior = str_c("Non-informative prior for intercept, mean = ", round(logit_scaled(our_belief2), digits = 2), ", sd = ", sd1) %>% str_wrap(width = 30)),
  tibble(log_odds = rnorm(1000, logit_scaled(our_belief2), sd2), prior = str_c("Non-informative prior for slope, mean = ", round(logit_scaled(our_belief2), digits = 2), ", sd = ", sd2) %>% str_wrap(width = 30)),
  tibble(log_odds = rnorm(1000, logit_scaled(our_belief2), sd3), prior = str_c("Weakly-informative prior for intercept, mean = ", round(logit_scaled(our_belief2), digits = 2), ", sd = ", sd3) %>% str_wrap(width = 30)),
  tibble(log_odds = rnorm(1000, logit_scaled(our_belief2), sd4), prior = str_c("Weakly-informative prior for slope, mean = ", round(logit_scaled(our_belief2), digits = 2), ", sd = ", sd4) %>% str_wrap(width = 30))) %>% mutate(probability = inv_logit_scaled(log_odds), prior = as_factor(prior))

prior_data %>% ggplot(aes(x = probability))+
  geom_density(fill = "grey80")+
  scale_x_continuous(limits = c(0,1), labels = c(0, "", 0.5, "", 1))+
  theme_classic()+
  facet_wrap(~prior, nrow = 2, ncol = 2, scales = "free")
```


Let's define a prior in log-odds scale

```{r}
our_prior2 = c(
  prior(normal(0,1.1), class = b),
  prior(normal(0,1.1), class = Intercept))

our_prior2 %>% as_tibble()
```


Let's run a model

```{r}
m5.4 = brm(formula = answer ~ sex, 
         data = data %>% mutate(answer = fct_relevel(answer, "Dog-person")), 
         family = bernoulli(), 
         file = "m5.4", 
         seed = 123,
         prior = our_prior2)
```



Model summary

```{r}
m5.4
```

Let's transform the intercept and slope

```{r}
#intercept: proportion of female students liking cats
inv_logit_scaled(-0.19)

#slope+intercept: proportion of male students liking cats
inv_logit_scaled(-0.19-0.68)
```

Let's plot our model: 
- females like cats and dogs
- males less likely like cats

```{r}
m5.4 %>% conditional_effects() 
```


Making predictions

```{r}
newdata = data.frame(sex = c("Female", "Male"))

m5.4 %>% fitted(m5.4, newdata = newdata)
```

Alternatives 

```{r}
#not back-transformed (log-odds) scale
m5.4 %>% posterior_linpred(newdata = newdata) %>% 
  as_tibble() %>%
  rename(Male = V1, Female = V2) %>% 
  posterior_summary() %>% 
  as_tibble(rownames = "Sex")

#transformed to probability scale
m5.4 %>% posterior_epred(newdata = newdata) %>% 
  as_tibble() %>%
  rename(Male = V1, Female = V2) %>% 
  posterior_summary() %>% 
  as_tibble(rownames = "Sex")

#transformed to probability scale (equivalent to the previous method)
m5.4 %>% posterior_linpred(newdata = newdata, transform = TRUE) %>% 
  as_tibble() %>%
  rename(Male = V1, Female = V2) %>% 
  posterior_summary() %>% 
  as_tibble(rownames = "Sex")
```

Let's now try answer a few study questions

First, let's make a transformed posterior

```{r}
posterior4 = m5.4 %>%
  as_draws_df() %>%
  inv_logit_scaled()
```


1) What is the probability that female students identify them as cat-persons?

hypothesis (brms)

```{r}
hyp1 = posterior4 %>% hypothesis("b_Intercept < 0.5")
hyp1
plot(hyp1)
```


bayestestR

```{r}
p_direction(posterior4$b_Intercept, null = 0.5)

pd1 = p_direction(posterior4$b_Intercept-0.5)

plot(pd1) + 
  scale_fill_discrete(name = "Answer", labels = c("Cat", "Dog"))+
  ggtitle("")
```


bayes R solution

```{r}
mean(posterior4$b_Intercept < 0.5)
```



2) What is the probability that male students identify them as cat-persons?

First, let's make a transformed posterior

```{r}
posterior5 = m5.4 %>%
  as_draws_df() %>%
  mutate(male_response = b_Intercept + b_sexMale) %>% 
  inv_logit_scaled() %>% 
  select(male_response, b_Intercept)
```


hypothesis (brms)

```{r}
hyp2 = posterior5 %>% hypothesis("male_response > 0.5")
hyp2
plot(hyp2)
```


bayestestR

```{r}
pd1 = p_direction(posterior5$male_response-0.5, null = 0.5)

plot(pd1) + 
  scale_fill_discrete(name = "Answer", labels = c("Dog", "Cat"))+
  ggtitle("")
```


bayes R solution

```{r}
mean(1-posterior5$male_response < 0.5)
```

3) How being a cat-person differs between sexes?

Effect size

```{r}
es = posterior5 %>% 
  rename("female_response" = b_Intercept) %>% 
  mutate(effect_size = female_response-male_response)

#difference in percentage points
es %>% posterior_summary()
```

Probability for females for identifying them as a cat person

```{r}
pd3 = p_direction(es$effect_size)

pd3

plot(pd3) + 
  scale_fill_discrete(name = "Probability for being a cat-person", labels = c("Male", "Female"))+
  ggtitle("")
```


How many times less likely males students identify them as cat-persons?

```{r}
posterior5 %>% 
  rename("female_response" = b_Intercept) %>% 
  mutate(difference_in_times = female_response/male_response) %>% 
  posterior_summary()
```


Odds ratio

```{r}
m5.4 %>% 
  as_draws_df() %>% 
  mutate(odds_ratio = exp(b_sexMale)) %>% 
  select(odds_ratio) %>% 
  posterior_summary() %>% 
  as_tibble(rownames = "Variable")
```


# Task on hip fracture data

Try to use the previous code to answer the following questions:

1) What is the probability for receiving post-discharge rehabilitation ("therapy" variable) after a hip fracture? Use an intercept-only model.

2) Do operatively and nonoperatively treated patients have equal opportunities for post-discharge rehabilitation? Use "treatment_method" as an X-variable.


## Use the following data

```{r}
set.seed(123)
all_HF_data = read_csv("data2022.csv")

#let's make the data smaller
HF_data = all_HF_data %>% 
  group_by(treatment_method) %>% 
  sample_frac(size = 0.08)

#Let's make a few necessary data manipulations
HF_data = HF_data %>% 
#1) let's use "nonoperative treatment" as a reference
    mutate(treatment_method = as_factor(treatment_method),
         treatment_method = fct_relevel(treatment_method, "operative"),
#2) make therapy received in phase II as a binary variable
therapy = case_when(therapy_phase_II == 0 ~ "Did not receive",
                             therapy_phase_II > 0 ~ "Received")) %>% 
#3) and select only analytically relevant variables
select(treatment_method, therapy)
```

1) What is the probability for receiving post-discharge rehabilitation ("therapy" variable) after a hip fracture? Use an intercept-only model.

# Create the intercept-only model

```{r}
m5.5 = brm(formula = answer ~ 1, 
         data = HF_data, 
         family = bernoulli(), 
         file = "m5.2", 
         seed = 123,
         sample_prior = "yes")
```

```{r}
m5.3
```

-- to transform intercept to a probability scale, we have to use inv_logit_scaled() function
```{r}
inv_logit_scaled(-0.16)
```

