---
title: "Practical session 2: the basics of regression"
output: html_notebook
---

# Let's prepare R for a new analysis

- Clear workspace
- Restart R


# If you haven't installed tidyverse, please do it now.

```{r}
install.packages(tidyverse)
```

# Load necessary package

```{r}
library(tidyverse)
```



# Let's simulate some data

```{r}

hippodata <- tibble(speed= c(20, 25, 32, 37, 34, 39, 36, 44, 48, 37, 28, 35, 32, 36, 38, 43, 41, 48, 45, 52),
                    tail_length = c(15, 12, 18, 14, 24, 22, 19, 30, 26, 32, 17, 16, 20, 22, 26, 25, 28, 28, 30, 32),
                    sex= rep(c(0, 1), each=10))



```


```{r eval=FALSE}
hippodata = bind_rows(
  tibble(
    speed = c(20, 25, 32, 37, 34, 39, 36, 44, 48, 37), 
    sex = c("Female"),
    tail_length = c(15, 12, 18, 14, 24, 22, 19, 30, 26, 32)), 
  tibble(
    speed = c(28, 35, 32, 36, 38, 43, 41, 48, 45, 52),
    sex = c("Male"),
    tail_length = c(17, 16, 20, 22, 26, 25, 28, 28, 30, 32)))
```

Let's view our data

```{r}
hippodata
```

# Average running speed in population

## EDA

Summary table

```{r}
hippodata %>% 
  summarise(mean = mean(speed),
            sd = sd(speed))
```

## Model

```{r}
m1 = lm(speed ~ 1, data = hippodata)

summary(m1)
```

Plot your model

```{r}
plot(data = hippodata,
        speed ~ 1)
abline(m1)
```


# Running speed difference between sexes

## EDA

Summary table

```{r}
hippodata %>% 
  group_by(sex) %>% 
  summarise(mean = mean(speed),
            sd = sd(speed))
```


Plot

```{r}
hippodata %>% 
  ggplot(aes(sex, speed))+
  geom_point()
```
## Model

```{r}
m2 = lm(speed ~ sex, data = hippodata)

summary(m2)
```

Plot your model

```{r}
ggplot(hippodata, aes(sex, speed)) + geom_point() + geom_abline(aes(slope= 4.6, intercept=35.2))
```


```{r eval=FALSE}
plot(data = hippodata,
        speed ~ sex)
abline(m2)
```



# Effect of tail length on running speed

## EDA

```{r}
hippodata %>% 
  ggplot(aes(tail_length, speed))+
  geom_point()
```


## Model

```{r}
m3 = lm(speed ~ tail_length, data = hippodata)

summary(m3)
```

Plot your model

```{r}
plot(data = hippodata,
        speed ~ tail_length)
abline(m3)
```

```{r}
plot(residuals(m3))
```



